The Spatial and Temporal Domains of Modern Ecology

### 2017-12-17 paper/main.tex
Removed cache statements in fig3, additional-analyses vignettes, some edits to calibration (remove commented out code), checking results in that. Edits in manuscript

\noindent \textbf{Results}\\
\noindent \textbf{\emph{Domains}}\\
In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B). Bar widths in (C-D) indicate differences in scale between x-axis labels.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists_bs.png}
\vspace{-0.15 cm}
\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. }
\label{afoto1}
\end{figure}

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig. 1D).

Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

\begin{figure}[ht]
\includegraphics[width=1\textwidth]{figures/act_v_eff_diff.png}
\vspace{-0.2 cm}
\caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B). Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.  The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.}
\label{afoto1}
\end{figure}

\noindent \textbf{\emph{Observational methods}}\\
We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each. Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance. Automated sensing methods showed no trend over time (SI).

@@

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

@@

>>>>

\noindent \textbf{Results}\\
\noindent \textbf{\emph{Observational methods}}\\
To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.9\%, and paleo-reconstruction for less than 0.8\%.

\noindent \textbf{\emph{Domains}}\\
In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B).

@@

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. Duration was one day or less for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig. 1D).

Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. Different observational methods occupied substantially different portions of the domain space, as indicated by the locations of the median resolution and interval values for each:  field observations had a resolution of 0.1-1 m$^2$ and were repeated at monthly intervals, whereas remote observations had coarser resolutions (1000 m$^2$) but finer repeat intervals ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily intervals while paleo-reconstructions had multi-decadal intervals.  The complete domain space of each observational type is detailed in the SI.

This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

%\noindent \textbf{\emph{Observational methods}}\\
%We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each.

@@

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.

Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance. Automated sensing methods showed no trend over time (SI).

Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{../vignettes/figures/fig1.png}
\vspace{-0.15 cm}
\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. Bar widths in C-D indicate differences in scale between x-axis labels. The grey vertical line in D indicates that the majority ($>$95\%) of observations of $\leq$1 day duration were temporally unreplicated.}
\label{afoto1}
\end{figure}

\begin{figure}[ht]
\includegraphics[width=1\textwidth]{../vignettes/figures/fig3.png}
\vspace{-0.2 cm}
\caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B). Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.  The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker. The grey vertical line in D indicates that the majority ($>$95\%) of observations of $\leq$1 day duration were temporally unreplicated.}
\label{afoto1}
\end{figure}
